---
title: "Exercise 4"
author: "Tobias Famos"
output: pdf_document
---
## Task 1
Load the data
```{r}
education <- read.table("EducationBis.txt", header=T)
education_men <- education[which(education$Gender== "male"), ]
edcuation_women <- education[which(education$Gender=="female"),]
```

Now build linear model to explain `Wage` by `Edcuation`
```{r}
linear_model_men <- lm(education_men$Wage~education_men$Education)
linear_model_women <- lm(edcuation_women$Wage~edcuation_women$Education)
summary(linear_model_men)
summary(linear_model_women)
```
For both models, the multiple R^2 is quite high, thus we can conclude that the linear model explains much of the variance.
The model for the women shows a significant p value in both the intercept and the slope, thus we can accpet it as true.
The model for the men has only a significant p value in the slope but not in the intercept.


## Task 2
The slopes are significantly different from 0 as they both have a p value < 0.5%.

## Task 3
First remove the ID column as we don't need it
```{r}
education_no_index <- subset(education, select=-c(ID))

```
```{r}
unifiedModel <- lm(education_no_index$Wage~., data=education_no_index)
summary(unifiedModel)
```

The unified model explains the Wage quite good. All variables are significant, and the R^2 is almost 1.

## Task 4
We can't use vendor name and model name to explain the performance.

## Task 5
```{r}
computers <- read.table("Computers.txt", header=T)
computers_no_vendor_model <- subset(computers, select=-c(vendor, model))

computeds_model <- lm(computers_no_vendor_model$PRP~. , data=computers_no_vendor_model)
summary(computeds_model)
```
From the preliminary linear regression, the single variable I would use for a linear regression is the `ERP`.

```{r}
singleModel <- lm(computers$PRP ~ computers$ERP)
summary(singleModel)
confint(singleModel)
```
The model explains the 93% of the variance, and it is quite significant.
The confidence interval is 2 times the standard error, thus 2*0.018 = 0.036
## Task 6
```{r}
plot(computers$PRP, computers$ERP, xlab="Published Relative Performance", ylab="Performance", main="Performance vs published realtive Performance " )
abline(singleModel$coefficients)
```

# Task 7
```{r}
cars = read.table("Cars.txt", header=T)
cars = subset(cars, select = -c(name))
all_model <- lm(cars$mpg~.,data=cars)
summary(all_model)
```
The not significant variables are : `cylinders`, `horsepower` and `acceleration`