---
title: "Exercise 4"
author: "Tobias Famos"
output: pdf_document
---
# Exercise 4
## Task 1
Load the data
```{r}
education <- read.table("EducationBis.txt", header=T)
education$Gender
education_men <- education[which(education$Gender== "male"), ]
edcuation_women <- education[which(education$Gender=="female"),]
```

Now build linear model to explain `Wage` by `Edcuation`
```{r}
linear_model_men <- lm(education_men$Wage~education_men$Education)
linear_model_women <- lm(edcuation_women$Wage~edcuation_women$Education)
summary(linear_model_men)
summary(linear_model_women)
```
For both models, the multiple R^2 is quite high, thus we can conclude that the linear model explains much of the variance.
The model for the women shows a significant p value in both the intercept and the slope, thus we can accpet it as true.
The model for the men has only a significant p value in the slope but not in the intercept.


## Task 2
The slopes are significantly different from 0 as they both have a p value < 0.5%.

## Task 3
First remove the ID column as we don't need it
```{r}
education_no_index <- subset(education, select=-c(ID))

```
```{r}
unifiedModel <- lm(education_no_index$Wage~., data=education_no_index)
summary(unifiedModel)
```

The unified model explains the Wage quite good. All variables are significant, and the R^2 is almost 1.

## Task 4