---
title: "Exercise 2"
author: "Tobias Famos"
output: pdf_document
---
## Reading the Data
```{r}
educationDataSet <- read.table("Education.txt", header=T)
```
## Exercise 1: Plotting
Yes you can change all the Labels on a graph with plot by adding `main`, `ylab` and `xlab` parameters.

```{r}
plot(educationDataSet$Education, educationDataSet$Wage, xlab="Education in Years", ylab="Wage per Year", main="Education vs Wage", xlim=c(0, max(educationDataSet$Education)))
```
 ## Filtering by Gender
First create the subsets
```{r}
males <- subset(educationDataSet, Gender == 1)
females <- subset(educationDataSet, Gender == 2)
```
Now plot the females and the males
```{r}
plot(females$Education, females$Wage, xlab="Education in Years", ylab="Wage per Year", main="Education vs Wage only Female", xlim=c(0, max(educationDataSet$Education)))
plot(males$Education, males$Wage, xlab="Education in Years", ylab="Wage per Year", main="Education vs Wage only Male", xlim=c(0, max(educationDataSet$Education)))
```
Also Plot them both with different colors.
```{r}
colors <- educationDataSet$Gender
plot(educationDataSet$Education, educationDataSet$Wage, xlab="Education in Years", ylab="Wage per Year", main="Education vs Wage", xlim=c(0, max(educationDataSet$Education)), col=colors)
```
## Sinmpson's Hostpial
```{r}
hospitalDataSet <- read.table("Simpson.txt", header=T)
hospitalDataSet$Treatment <- factor(hospitalDataSet$Treatment)
drugs <- subset(hospitalDataSet, as.integer(Treatment)==1)
surgery <- subset(hospitalDataSet, as.integer(Treatment)==2)
summary(drugs)
summary(surgery)
```
Just from loking at the mean of the Result, the Drugs seem to be more effective, as the drug mean is 1.239 where the Treatment mean is 1342
```{r}
drugsSmall <- subset(drugs, Size == 1)
surgerySmall <- subset(surgery, Size == 1)
summary(drugsSmall)
summary(surgerySmall)
```
Yes I change my recommondation as for the small tumors, ther surgery has a mean of 1.105 and the drugs have a mean of the result of 1.18.


If the tumor is large:
```{r}
drugLarge <- subset(drugs, Size == 2)
surgeryLarge <- subset(surgery, Size == 2)
summary(drugLarge)
summary(surgeryLarge)
```
I would recommend the Surgery, as it has a lower mean on the Result. Thus is more effective.


## Am I Consistent?

I am not consistent in the cases above.
The treatments have different effectivenesses for the size of the tumors.
Thus the treatment decision must be done by taking into account the size of the tumor
