---
title: "Exercise 3"
author: "Tobias Famos"
output: pdf_document
---
## Reading the Data
```{r}
original_data <- read.table("Mean20.txt", header=T)
```
Computing the Median, Mean, Standard Deviation, min and maximum of the variable time.
```{r}
summary(original_data)
```
As we can see with the summary command, we need to do some preprocessing of the dataset, as there is a negative time and one NA.
```{r}
meanDataSet <- na.omit(original_data)
meanDataSet <- meanDataSet[which(meanDataSet$time >=0),]
summary(meanDataSet)
plot(meanDataSet)
```
Calculate the Min, Max, mean, standard deviation and median
one NA.
```{r}
media <- median(meanDataSet)
mean <- mean(meanDataSet)
min <- min(meanDataSet)
max <- max(meanDataSet)
std <- sd(meanDataSet)
```
## Exercise 2
```{r}
t.test(meanDataSet, mu=7.05)
t.test(original_data, mu=7.05)
```
We can find no significant indication that the mean is 7.05. Thus the alternative hypothesis must be concidered true, that the mean is not equal to 7.05

The difference to the original Data set is that the confidence interval is larger.

## Question 3
With this knowledge we can do a one sided test.
```{r}
t.test(meanDataSet, mu=7.05, alternative = "greater")
```
Same thing, it is still not significant.

# Qeustion 4
```{r}
secondMax <- function (vector){
  if(!is.vector(vector)){
    stop("Argument must be a vector")
  }
  n <- length(vector)
  return( sort(vector,partial=n-1)[n-1])
}
secondMax(meanDataSet)
secondMax(c(1,2,3,4))

```
## Question 5

I think the Goal is to return the Summary so I will do this.
```{r}
mySummary <- function(x){
  media <- median(meanDataSet)
  mean <- mean(meanDataSet)
  min <- min(meanDataSet)
  max <- max(meanDataSet)
  std <- sd(meanDataSet)
  return(c(mean, media, std, min, max))
}

mySummary(medianDataSet)
```
