---
title: "Statistical Learning Methods Exercise 11"
output: pdf_document
author: "Tobias Famos"
---
# Preliminaries

Loading the Votes
```{r}
votes <- read.table("/home/tobias/unibe/statistical methods in R/Exercise 11/Votes.txt", header=T)
sum(is.na(votes))
```

Add the row names to prevent objects to only show up as numbers in plots
```{r}
rownames(votes) <- votes$Vote

```
# Task 1 Cluster the votes from 2000 until 20145
Select the votes for the clustering
```{r}
votes_00_to_14 <- votes[28:161]
dist_00_to_14 <- dist(votes_00_to_14)
cluster_00_to_14 <- hclust(dist_00_to_14, method="complete")
plot(cluster_00_to_14)
```

# Task 2 Cluster votes from 2014 to 2017

```{r}
votes_14_to_17 <- votes[0:27]
dist_14_to_17 <- dist(votes_14_to_17)
labels <- votes$Vote
cluster_14_to_17 <- hclust(dist_14_to_17, method="complete")
plot(cluster_14_to_17)
```

# Task 3 Compare
```{r}
library(dendextend)
dend00 <- as.dendrogram (cluster_00_to_14)
dend14 <- as.dendrogram (cluster_14_to_17)

dend_list <- dendlist(dend00, dend14)

dendlist(dend00, dend14) %>%
  untangle(method = "step1side") %>% # Find the best alignment layout
  tanglegram()

```
In both clustering approaches there are two pretty distinct clusters emerging.
When comparing the two pretty distinct clusters, the only different canton in them is the TI.

# Task 4 Average Link method
For 2000 to 2014
```{r}
cluster_00_to_14_average <- hclust(dist_00_to_14, method="average")
plot(cluster_00_to_14_average)
```
```{r}
dend <- as.dendrogram (cluster_00_to_14)
dendAverage <- as.dendrogram (cluster_00_to_14_average)

dend_list <- dendlist(dend, dendAverage)

dendlist(dend, dendAverage) %>%
  untangle(method = "step1side") %>% # Find the best alignment layout
  tanglegram()
```

When clustering the same data with the average method, we get again two pretty distinct clusters, but the members of
the clusters have changed quite a lot


```{r}
cluster_14_to_17_average <- hclust(dist_14_to_17, method="average")
plot(cluster_14_to_17_average)
```
```{r}
dend <- as.dendrogram (cluster_14_to_17)
dendAverage <- as.dendrogram (cluster_14_to_17_average)

dend_list <- dendlist(dend, dendAverage)

dendlist(dend, dendAverage) %>%
  untangle(method = "step1side") %>% # Find the best alignment layout
  tanglegram()
```
The same thing holds for the 2014 to 2017 Data